---
title: "Allan Zhao_A2"
author: "Allan Zhao"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, messaage = F)

library(tidyverse)
library(sf)
library(mapview)
library(censusapi)
library(tigris)
library(s2)
library(leaflet)
```

### Setup for pulling census data:
```{r}
Sys.setenv(CENSUS_KEY="eb8ddadea31c0afe47f0aa2e05a42c91a84439fe")
```

##Showing general trends in population change:
### Importing and combining general population data for 2020 and 2010 
### censuses for Santa Clara County
```{r}
scc_pop_2020 <-
  getCensus(
    name = "dec/pl",
    vintage = 2020,
    region = "block:*", 
    regionin = "state:06+county:085",
    vars = "P1_001N"
  ) %>% 
  transmute(
    block =
      paste0(state,county,tract,block),
    pop = P1_001N
  )
colnames(scc_pop_2020) <- c("block", "pop20")

scc_pop_2010 <-
  getCensus(
    name = "dec/pl",
    vintage = 2010,
    region = "block:*", 
    regionin = "state:06+county:085",
    vars = "P003001"
  ) %>% 
  transmute(
    block =
      paste0(state,county,tract,block),
    pop = P003001
  )
colnames(scc_pop_2010) <- c("block", "pop10")
```

### Manip subsets
```{r}
scc_blocks_2020 <- blocks("CA", "Santa Clara", year = 2020, progress_bar = F)
scc_blocks_2010 <- blocks("CA", "Santa Clara", year = 2010, progress_bar = F)

scc_pop_blocks <- scc_blocks_2020 %>% 
  select(block = GEOID20) %>% 
  left_join(scc_pop_2020)
```

Manip 2020 pop blocks to 2010 boundaries
```{r}
scc_pop_blocks_1 <-
  scc_pop_blocks %>% 
  st_centroid() %>% 
  .[scc_blocks_2010, ] %>% 
  st_set_geometry(NULL) %>% 
  left_join(scc_pop_blocks %>% select(block)) %>% 
  st_as_sf()

sum(scc_pop_blocks$pop20)
sum(scc_pop_blocks_1$pop20)
```

Transforming block areas
```{r}
scc_pop_blocks_area <-
  scc_pop_blocks %>% 
  st_transform(26910) %>% 
  mutate(area = st_area(.))
```

Running intersection
```{r}
scc_pop_blocks_intersection <-
  scc_pop_blocks_area %>% 
  st_intersection(
    scc_blocks_2010 %>% 
      st_transform(26910)
  )
```

Final proportional manip of population to new boundaries
```{r}
scc_pop_blocks_final <-
  scc_pop_blocks_1 %>% 
  select(block = geometry) %>% 
  left_join(scc_blocks_2020) %>% 
  st_transform(26910) %>% 
  mutate(original_area = st_area(.)) %>% 
  st_intersection(
    scc_blocks_2010 %>% 
      st_transform(26910)
  ) %>% 
  mutate(
    leftover_area = st_area(.),
    perc_area = leftover_area / original_area,
    pop = pop * perc_area
  )
```

```{r}
scc_pop <- merge(scc_pop_blocks_1, scc_pop_2010, by.y = "block")

scc_pop$pop_change <- scc_pop$pop20 - scc_pop$pop10
```

### Mapping 2020 population
```{r}
mlpts_boundary <- places("CA", progress_bar = F) %>% 
  filter(NAME == "Milpitas")

mlpts_pop <- scc_pop %>% 
  left_join(scc_blocks_2010 %>% select(block = geometry)) %>% 
  st_as_sf() %>% 
  st_centroid() %>% 
  .[mlpts_boundary, ] %>% 
  st_set_geometry(NULL) %>% 
  left_join(scc_blocks_2010 %>% select(block = geometry)) %>% 
  st_as_sf()

mapview(mlpts_pop, zcol = "pop_change")
```

----
## Analyzing data by race:
### Defining decennial census variables
```{r}
dec_vars_2020 <-
  listCensusMetadata(
    name = "2020/dec/pl",
    type = "variables"
  )

dec_vars_2010 <-
  listCensusMetadata(
    name = "2010/dec/pl",
    type = "variables"
  )
```

### Drawing 2020 decennial census data for Santa Clara County
```{r}
scc_pop_race_2020 <-
  getCensus(
    name = "dec/pl",
    vintage = 2020,
    region = "block:*", 
    regionin = "state:06+county:085",
    vars = "group(P1)"
  ) %>% 
  mutate(
    block =
      paste0(state,county,tract,block)
  ) %>% 
  select(!c(GEO_ID,state,county,tract,NAME) & !ends_with(c("NA"))) %>% 
  pivot_longer(
    ends_with("N"),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    dec_vars_2020 %>% 
      select(name, label)
  ) %>% 
  select(-name) %>% 
  separate(
    label,
    into = c(NA,NA,"category1","category2"),
    sep = "!!"
  )
```

### Drawing 2010 decennial data for Santa Clara County  [[SOMETHING DOESN'T WORK]]
```{r}
scc_pop_race_2010 <-
  getCensus(
    name = "dec/pl",
    vintage = 2010,
    region = "block:*", 
    regionin = "state:06+county:085",
    vars = "group(P1)"
  ) %>% 
  mutate(
    block =
      paste0(state,county,tract,block)
  ) %>% 
  select(!c(GEO_ID,state,county,tract) & !ends_with(c("NA"))) %>% 
  pivot_longer(
    ends_with("N"),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    dec_vars_2010 %>% 
      select(name, label)
  ) %>% 
  select(-name) %>% 
  separate(
    label,
    into = c(NA,NA,"category1","category2"),
    sep = "!!"
  )
```


### Separating above dataset into race and cleaning data.
```{r}
scc_pop_race_2020 <- scc_pop_race_2020 %>% 
  mutate(
    race = case_when(
      category1 == "Population of two or more races:" & is.na(category2) ~ "Two or more races",
      category1 == "Population of two or more races:" ~ "",
      !is.na(category2) ~ category2,
      TRUE ~ ""
    )
  )

scc_pop_race_2020 <- scc_pop_race_2020 %>% 
  filter(race != "") %>% 
  select(block, race, pop = estimate)
```





























