---
title: "AllanZhao_A1"
author: "Allan Zhao"
date: "September 28, 2021"
output: html_document
---

## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(plotly)
```

## Loading PG&E data from CSVs
```{r}
## Loading 2021 electricity files
pge_21_q1_elec <- read_csv("PG&E/PGE_2021_Q1_ElectricUsageByZip.csv")
pge_21_q2_elec <- read_csv("PG&E/PGE_2021_Q2_ElectricUsageByZip.csv")

## Loading 2020 electricity files
pge_20_q1_elec <- read_csv("PG&E/PGE_2020_Q1_ElectricUsageByZip.csv")
pge_20_q2_elec <- read_csv("PG&E/PGE_2020_Q2_ElectricUsageByZip.csv")
pge_20_q3_elec <- read_csv("PG&E/PGE_2020_Q3_ElectricUsageByZip.csv")
pge_20_q4_elec <- read_csv("PG&E/PGE_2020_Q4_ElectricUsageByZip.csv")

## Loading 2019 electricity files
pge_19_q1_elec <- read_csv("PG&E/PGE_2019_Q1_ElectricUsageByZip.csv")
pge_19_q2_elec <- read_csv("PG&E/PGE_2019_Q2_ElectricUsageByZip.csv")
pge_19_q3_elec <- read_csv("PG&E/PGE_2019_Q3_ElectricUsageByZip.csv")
pge_19_q4_elec <- read_csv("PG&E/PGE_2019_Q4_ElectricUsageByZip.csv")

## Loading 2018 electricity files
pge_18_q1_elec <- read_csv("PG&E/PGE_2018_Q1_ElectricUsageByZip.csv")
pge_18_q2_elec <- read_csv("PG&E/PGE_2018_Q2_ElectricUsageByZip.csv")
pge_18_q3_elec <- read_csv("PG&E/PGE_2018_Q3_ElectricUsageByZip.csv")
pge_18_q4_elec <- read_csv("PG&E/PGE_2018_Q4_ElectricUsageByZip.csv")

## Loading 2017 electricity files
pge_17_q1_elec <- read_csv("PG&E/PGE_2017_Q1_ElectricUsageByZip.csv")
pge_17_q2_elec <- read_csv("PG&E/PGE_2017_Q2_ElectricUsageByZip.csv")
pge_17_q3_elec <- read_csv("PG&E/PGE_2017_Q3_ElectricUsageByZip.csv")
pge_17_q4_elec <- read_csv("PG&E/PGE_2017_Q4_ElectricUsageByZip.csv")


#####################################################################
#Loading 2021 gas files
pge_21_q1_gas <- read_csv("PG&E/PGE_2021_Q1_GasUsageByZip.csv")
pge_21_q2_gas <- read_csv("PG&E/PGE_2021_Q2_GasUsageByZip.csv")

## Loading 2020 electricity files
pge_20_q1_gas <- read_csv("PG&E/PGE_2020_Q1_GasUsageByZip.csv")
pge_20_q2_gas <- read_csv("PG&E/PGE_2020_Q2_GasUsageByZip.csv")
pge_20_q3_gas <- read_csv("PG&E/PGE_2020_Q3_GasUsageByZip.csv")
pge_20_q4_gas <- read_csv("PG&E/PGE_2020_Q4_GasUsageByZip.csv")

## Loading 2019 electricity files
pge_19_q1_gas <- read_csv("PG&E/PGE_2019_Q1_GasUsageByZip.csv")
pge_19_q2_gas <- read_csv("PG&E/PGE_2019_Q2_GasUsageByZip.csv")
pge_19_q3_gas <- read_csv("PG&E/PGE_2019_Q3_GasUsageByZip.csv")
pge_19_q4_gas <- read_csv("PG&E/PGE_2019_Q4_GasUsageByZip.csv")

## Loading 2018 electricity files
pge_18_q1_gas <- read_csv("PG&E/PGE_2018_Q1_GasUsageByZip.csv")
pge_18_q2_gas <- read_csv("PG&E/PGE_2018_Q2_GasUsageByZip.csv")
pge_18_q3_gas <- read_csv("PG&E/PGE_2018_Q3_GasUsageByZip.csv")
pge_18_q4_gas <- read_csv("PG&E/PGE_2018_Q4_GasUsageByZip.csv")

## Loading 2017 electricity files
pge_17_q1_gas <- read_csv("PG&E/PGE_2017_Q1_GasUsageByZip.csv")
pge_17_q2_gas <- read_csv("PG&E/PGE_2017_Q2_GasUsageByZip.csv")
pge_17_q3_gas <- read_csv("PG&E/PGE_2017_Q3_GasUsageByZip.csv")
pge_17_q4_gas <- read_csv("PG&E/PGE_2017_Q4_GasUsageByZip.csv")

```

## Saving data into RDS files by year
```{r}
# 2021 Electric
year <- 2021
quarters <- 1:2
type <- "Electric"

pge_21_elec <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PG&E/PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_21_elec <- rbind(pge_21_elec,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_21_elec, "pge_21_elec.rds")
}
```

```{r}
# 2020 Electric
year <- 2020
quarters <- 1:4
type <- "Electric"

pge_20_elec <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PG&E/PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_20_elec <- rbind(pge_20_elec,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_20_elec, "pge_20_elec.rds")
}
```

```{r}
# 2019 Electric
year <- 2019
quarters <- 1:4
type <- "Electric"

pge_19_elec <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PG&E/PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_19_elec <- rbind(pge_19_elec,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_19_elec, "pge_19_elec.rds")
}
```

```{r}
# 2018 Electric
year <- 2018
quarters <- 1:4
type <- "Electric"

pge_18_elec <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PG&E/PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_18_elec <- rbind(pge_18_elec,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_18_elec, "pge_18_elec.rds")
}
```

```{r}
# 2017 Electric
year <- 2017
quarters <- 1:4
type <- "Electric"

pge_17_elec <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PG&E/PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_17_elec <- rbind(pge_17_elec,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_17_elec, "pge_17_elec.rds")
}
```

```{r}
# 2021 Gas
year <- 2021
quarters <- 1:2
type <- "Gas"

pge_21_gas <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PG&E/PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_21_gas <- rbind(pge_21_gas,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_21_gas, "pge_21_gas.rds")
}
```

```{r}
# 2020 Gas
year <- 2020
quarters <- 1:4
type <- "Gas"

pge_20_gas <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PG&E/PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_20_gas <- rbind(pge_20_gas,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_20_gas, "pge_20_gas.rds")
}
```

```{r}
# 2019 Gas
year <- 2019
quarters <- 1:4
type <- "Gas"

pge_19_gas <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PG&E/PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_19_gas <- rbind(pge_19_gas,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_19_gas, "pge_19_gas.rds")
}
```

```{r}
#2018 Gas
year <- 2018
quarters <- 1:4
type <- "Gas"

pge_18_gas <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PG&E/PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_18_gas <- rbind(pge_18_gas,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_18_gas, "pge_18_gas.rds")
}
```

```{r}
# 2017 Gas
year <- 2017
quarters <- 1:4
type <- "Gas"

pge_17_gas <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PG&E/PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_17_gas <- rbind(pge_17_gas,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_17_gas, "pge_17_gas.rds")
}
```

## Creating dataframes with all years' data combined (for easier graphing manip)
```{r}
pge_elec_all <- rbind(pge_17_elec, 
                      pge_18_elec, 
                      pge_19_elec, 
                      pge_20_elec, 
                      pge_21_elec)

pge_elec_all

pge_gas_all <- rbind(pge_17_gas,
                     pge_18_gas,
                     pge_19_gas,
                     pge_20_gas,
                     pge_21_gas)

pge_gas_all
```

## Simplifying two aggregated dataframes via piped example template in Ch 1.7
```{r}
pge_elec_final <-
  pge_elec_all %>% 
  filter(
    CUSTOMERCLASS %in% 
      c(
        "Elec- Residential",
        "Elec- Commercial"
      )
  ) %>% 
  select(
    -c(COMBINED, AVERAGEKWH)
  ) %>% 
  group_by(YEAR, MONTH, CUSTOMERCLASS) %>% 
  summarize(
    TOTALKWH = 
      sum(
        TOTALKWH, 
        na.rm = T
      ),
    TOTALCUSTOMERS =
      sum(
        TOTALCUSTOMERS,
        na.rm = T
      )
  ) %>% 
  mutate(
    AVERAGEKWH =
      TOTALKWH/TOTALCUSTOMERS
  )

pge_elec_final
```

```{r}
pge_gas_final <-
  pge_gas_all %>% 
  filter(
    CUSTOMERCLASS %in% 
      c(
        "Gas- Residential",
        "Gas- Commercial"
      )
  ) %>% 
  select(
    -c(COMBINED, AVERAGETHM)
  ) %>% 
  group_by(YEAR, MONTH, CUSTOMERCLASS) %>% 
  summarize(
    TOTALTHM = 
      sum(
        TOTALTHM, 
        na.rm = T
      ),
    TOTALCUSTOMERS =
      sum(
        TOTALCUSTOMERS,
        na.rm = T
      )
  ) %>% 
  mutate(
    AVERAGETHM =
      TOTALTHM/TOTALCUSTOMERS
  )

pge_gas_final
```

## Concatenating month & year columns (for easier graphing manip)
```{r}
pge_elec_final$DATE <- as.Date(
                          paste(
                            pge_elec_final$YEAR, 
                            pge_elec_final$MONTH, 
                            "01",
                            sep = "-")
                        )

pge_gas_final$DATE <- as.Date(
                          paste(
                            pge_gas_final$YEAR, 
                            pge_gas_final$MONTH, 
                            "01",
                            sep = "-")
                        )

```

## Converting gas & electric to universal units
```{r}
pge_elec_final$TOTALKBTU <- pge_elec_final$TOTALKWH * 3.412
# 3.412 is the multiplier for a conversion from KWH to kBTU
# Source: https://sciencing.com/calculate-kilowatt-hours-4902973.html

pge_gas_final$TOTALKBTU <- pge_gas_final$TOTALTHM * 100
# 100 is the approx. multiplier for a convesion from therms to kBTU
# Source: https://sciencing.com/convert-ccf-mmbtu-2825.html
```

## Filtering remaining dataframes by customer type
```{r}
pge_elec_residential <- filter(
                          pge_elec_final, 
                          CUSTOMERCLASS == "Elec- Residential")
pge_elec_commercial <- filter(
                          pge_elec_final, 
                          CUSTOMERCLASS == "Elec- Commercial")

pge_gas_residential <- filter(
                          pge_gas_final, 
                          CUSTOMERCLASS == "Gas- Residential")
pge_gas_commercial <- filter(
                          pge_gas_final, 
                          CUSTOMERCLASS == "Gas- Commercial")
```

## Creating plots for Residential Gas & Electric (2017 - present)
```{r}
pge_res_chart <- 
    ggplot() +
    geom_line(
      aes(
       x = pge_elec_residential$DATE,
       y = pge_elec_residential$TOTALKBTU,
       fill = pge_elec_residential$CUSTOMERCLASS
     ),
     stat = "identity",
     position = "stack",
     color = "purple"
   ) +
    geom_line(
      aes(
       x = pge_gas_residential$DATE,
       y = pge_gas_residential$TOTALKBTU,
       fill = pge_gas_residential$CUSTOMERCLASS
     ),
     stat = "identity",
     position = "stack",
     color = "orange"
    ) +
    labs(
      x = "Month",
      y = "KBTU",
       title = "PG&E Territory Monthly Residential Gas & Electricity Usage 
      2017-Present (in kBTUs)",
      fill = "Customer Type"
  )

pge_res_chart %>% ggplotly()
```

## Creating plots for Commercial Gas & Electric (2017 - present)
```{r}
pge_comm_chart <- 
    ggplot() +
    geom_line(
      aes(
       x = pge_elec_commercial$DATE,
       y = pge_elec_commercial$TOTALKBTU,
       fill = pge_elec_commercial$CUSTOMERCLASS
     ),
     stat = "identity",
     position = "stack",
     color = "purple"
   ) +
    geom_line(
      aes(
       x = pge_gas_commercial$DATE,
       y = pge_gas_commercial$TOTALKBTU,
       fill = pge_gas_commercial$CUSTOMERCLASS
     ),
     stat = "identity",
     position = "stack",
     color = "orange"
    ) +
    labs(
      x = "Month",
      y = "KBTU",
       title = "PG&E Territory Monthly Commercial Gas & Electricity Usage 
      2017-Present (in kBTUs)",
      fill = "Customer Type"
  )

pge_comm_chart %>% ggplotly()
```

## Observations:

### On residential customers:
Electricity usage by residential customers appears to have increased since
the onset of COVID. This can be explained by differences in patterns of behavior
once moving indoors. However, this may not capture the entire picture of 
electricity usage. The rise of Community Choice Aggregators (CCAs) within PG&E's 
service territory has increased dramatically since 2018, taking away portions
of PG&E's electricity load that would otherwise be represented.

Gas usage seems to be within general trends (perhappps, slightly higher than usual),
and with a marked anomaly around March 2020. 

## On commercial customers:
There is a significant decrease in electricity usage by commercial customers in 
2020, presumably because of COVID. This is evident by the precipitous drop in 
usage from March 2020 onwards. This trend continues until Q2 of 2021, where 
usage begins to climb again.

There is a slight decline in commercial gas usage in 2020, but not as dramatic
as the change with electricity usage. Perhaps this discrepancy can once again
be accounted for by the shifting of electric load to CCAs and away from PG&E 
(CCAs do not sell gas).




















