---
title: "Assignment 1"
author: "Allan Zhao"
date: "9.28.2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading tidyverse and plotly package through library() command
```{r}
library(tidyverse)
library(plotly)
```

# Loading all PG&E data from CSVs
```{r}
## Loading 2021 files
pge_21_q1_elec <- read_csv("PGE_2021_Q1_ElectricUsageByZip.csv")
pge_21_q2_elec <- read_csv("PGE_2021_Q2_ElectricUsageByZip.csv")

## Loading 2020 files
pge_20_q1_elec <- read_csv("PGE_2020_Q1_ElectricUsageByZip.csv")
pge_20_q2_elec <- read_csv("PGE_2020_Q2_ElectricUsageByZip.csv")
pge_20_q3_elec <- read_csv("PGE_2020_Q3_ElectricUsageByZip.csv")
pge_20_q4_elec <- read_csv("PGE_2020_Q4_ElectricUsageByZip.csv")

## Loading 2019 files
pge_19_q1_elec <- read_csv("PGE_2019_Q1_ElectricUsageByZip.csv")
pge_19_q2_elec <- read_csv("PGE_2019_Q2_ElectricUsageByZip.csv")
pge_19_q3_elec <- read_csv("PGE_2019_Q3_ElectricUsageByZip.csv")
pge_19_q4_elec <- read_csv("PGE_2019_Q4_ElectricUsageByZip.csv")

## Loading 2018 files
pge_18_q1_elec <- read_csv("PGE_2018_Q1_ElectricUsageByZip.csv")
pge_18_q2_elec <- read_csv("PGE_2018_Q2_ElectricUsageByZip.csv")
pge_18_q3_elec <- read_csv("PGE_2018_Q3_ElectricUsageByZip.csv")
pge_18_q4_elec <- read_csv("PGE_2018_Q4_ElectricUsageByZip.csv")

## Loading 2017 files
pge_17_q1_elec <- read_csv("PGE_2017_Q1_ElectricUsageByZip.csv")
pge_17_q2_elec <- read_csv("PGE_2017_Q2_ElectricUsageByZip.csv")
pge_17_q3_elec <- read_csv("PGE_2017_Q3_ElectricUsageByZip.csv")
pge_17_q4_elec <- read_csv("PGE_2017_Q4_ElectricUsageByZip.csv")

```

# Saving 2021 data into RDS file by year
```{r}
year <- 2021
quarters <- 1:2
type <- "Electric"

pge_21_elec <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_21_elec <- rbind(pge_21_elec,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_21_elec, "pge_21_elec.rds")
}
```

# Saving 2020 data into RDS file by year
```{r}
year <- 2020
quarters <- 1:4
type <- "Electric"

pge_20_elec <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_20_elec <- rbind(pge_20_elec,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_20_elec, "pge_20_elec.rds")
}
```

# Saving 2019 data into RDS file by year
```{r}
year <- 2019
quarters <- 1:4
type <- "Electric"

pge_19_elec <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_19_elec <- rbind(pge_19_elec,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_19_elec, "pge_19_elec.rds")
}
```

# Saving 2018 data into RDS file by year
```{r}
year <- 2018
quarters <- 1:4
type <- "Electric"

pge_18_elec <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_18_elec <- rbind(pge_18_elec,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_18_elec, "pge_18_elec.rds")
}
```

# Saving 2017 data into RDS file by year
```{r}
year <- 2017
quarters <- 1:4
type <- "Electric"

pge_17_elec <- NULL

for(quarter in quarters) {
  
  filename <- 
    paste0(
      "PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  
  temp <- read_csv(filename)
  
  pge_17_elec <- rbind(pge_17_elec,temp)
  # Note rbind requires field names to be consistent for every new thing that you add.

  saveRDS(pge_17_elec, "pge_17_elec.rds")
}
```

# Creating dataframe with all years' data combined
```{r}
pge_all_years <- rbind(pge_17_elec, 
                       pge_18_elec, 
                       pge_19_elec, 
                       pge_20_elec, 
                       pge_21_elec)

pge_all_years
```

# Simplifying pge_all_years via piped example template in Ch 1.7
```{r}
pge_final <-
  pge_all_years %>% 
  filter(
    CUSTOMERCLASS %in% 
      c(
        "Elec- Residential",
        "Elec- Commercial"
      )
  ) %>% 
  select(
    -c(COMBINED, AVERAGEKWH)
  ) %>% 
  group_by(YEAR, MONTH, CUSTOMERCLASS) %>% 
  summarize(
    TOTALKWH = 
      sum(
        TOTALKWH, 
        na.rm = T
      ),
    TOTALCUSTOMERS =
      sum(
        TOTALCUSTOMERS,
        na.rm = T
      )
  ) %>% 
  mutate(
    AVERAGEKWH =
      TOTALKWH/TOTALCUSTOMERS
  )

pge_final
```

# Separating pge_all_years into two dataframes based on customer class
```{r}
pge_residential <- filter(pge_final, CUSTOMERCLASS == "Elec- Residential")

pge_commercial <- filter(pge_final, CUSTOMERCLASS == "Elec- Commercial")
```

```{r}
pge_res_chart <- 
    pge_residential %>%
    ggplot() +
    geom_bar(
      aes(
       x = YEAR %>% factor(),
       y = TOTALKWH,
       fill = CUSTOMERCLASS
     ),
     stat = "identity",
     position = "stack"
   ) +
    labs(
      x = "Year",
      y = "kWh",
       title = "PG&E Territory Yearly Electricity Usage, 2017-Present",
      fill = "Electricity Type"
  )

pge_res_chart %>% ggplotly()

pge_res_chart
```

